Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-19 doc/manuscript/translocation.qmd
Minor edits throughout based on final review

One consequence of human alteration of the biosphere is an increase in emerging infectious diseases [@jones2008; @fisher2012].
Such diseases pose a severe threat to wildlife populations [@daszak2000], and have caused dramatic declines and extinctions in a wide range of taxa, including echinoderms, mammals, birds, and amphibians [@hewson2014; @samuel2015; @scheele2019; @cunningham2021].
Amid this global loss of biodiversity, some populations decimated by disease have subsequently recovered to varying extents [@newell2013; @voyles2018; @knapp2016].
Although we know relatively little about the mechanisms underlying recovery [@brannelly2021; @russell2020], host evolution is one possibility [@carlson2014; @searle2020].
This process is known from a diversity of wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], and often results in increased host resistance and/or tolerance (see [@medzhitov2012] for definitions).
However, linking these evolutionary responses to population recoveries is difficult and few clear examples exist.
In addition to promoting recovery of disease-impacted populations, the resistant/tolerant individuals in such populations could potentially be used to reestablish populations that were extirpated following disease emergence [@mendelson2019; @brannelly2021].
To date, the use of resistant/tolerant individuals for species recovery is rare [but see @joseph2018], and studies that evaluate its potential utility are critically needed.

@@

Results from the individual variant and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C).
One of the 7 identified genes (LOC108802036) did not have an associated annotation.
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong - and parallel - selection at this locus associated with exposure of frogs to Bd across multiple populations.
(See SI Appendix for additional description of these results.)

The splined window analysis identified 33 outlier regions for $\pi_{diff}$ and 58 outlier regions for *F~ST~* (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics ("shared regions") and 2 of these shared regions also contained one or more outlier SNPs (described above).
A total of 33 annotated genes were found in the 9 shared regions (SI Appendix: Datasets S3, S4).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a large negative $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended major histocompatibility complex (MHC) Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether frogs from recovering populations could be used to reestablish extirpated populations, we conducted `r n_distinct(d1$site_id, d1$date)` reintroductions in Yosemite National Park (`r min(d1$year)`--`r max(d1$year)`).
Each of the reintroductions involved collection of adult frogs from 1 of `r n_distinct(d1$donor)` recovering, Bd-positive "donor" populations and translocating them to 1 of `r n_distinct(d1$site_id)` nearby recipient sites (@fig-yosemap) [@seddon2014].
The donor populations included 2 of the 5 recovering populations in the frog evolution study described above (@fig-allelefrequencies A: population 1 and 4).
Following translocation, adult survival and recruitment of new adults were estimated from CMR surveys and counts of tadpoles and juveniles were obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1-year period following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate), which we suggest resulted primarily from our improved ability to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (@fig-bdload-beforeafter, see SI Appendix text for additional details) [@joseph2018; @vredenburg2010].

Of the `r nrow(d2)` translocated populations, `r frog_counts_gt0_nrow` (`r frog_counts_gt0_nrow / nrow(d2)`) showed evidence of successful reproduction in subsequent years, as indicated by the presence of tadpoles and/or juveniles.
For these `r frog_counts_gt0_nrow` populations, one or both life stages were detected in nearly all survey-years following translocation (proportion of survey-years: median = `r frog_counts_gt0_median`, range = `r frog_counts_gt0_min`--`r frog_counts_gt0_max`).
These same populations were also those in which recruitment of new adults (i.e., progeny of translocated individuals) was detected.
As with early life stages, recruits were detected in the majority of post-translocation survey-years (proportion of survey-years: median = `r frog_captures_gt0_median`, range = `r frog_captures_gt0_min`--`r frog_captures_gt0_max`).
In summary, survey results suggest that translocations resulted in the establishment of reproducing MYL frog populations at most recipient sites despite the ongoing presence of Bd.

To determine whether Bd load had a negative effect on adult survival, as would be expected if frogs were highly susceptible to Bd infection, we developed linear and multilevel models.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects, @fig-survival-postdens).
Collectively, these results are consistent with frogs in recovering populations having evolved sufficient resistance/tolerance to minimize the adverse effects of Bd infection.
In addition, the moderate Bd loads that characterized frogs following translocation (@fig-bdload-beforeafter) and relatively high survival of frogs in numerous translocated cohorts (@fig-translocation-survival) suggests that frogs translocated from recovering populations can maintain the benefits of resistance/tolerance in non-natal habitats.
Finally, in 3 locations where longer CMR time series allowed us to assess the survival of new adults recruited to the population, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs).
Therefore, frog resistance/tolerance in non-natal habitats is maintained across generations.

@@

Even when incorporating (i) demographic stochasticity and (ii) environmental stochasticity in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental variability in the presence of Bd), populations with high adult survival are likely to persist over a 50 year time horizon.
Our model predicted that, following a single introduction of 40 adult individuals into a population and using the baseline value for $\omega$ of 0.3, the six populations with the highest adult survival probabilities had 50-year extinction probabilities of 0 to 0.32 (@fig-viability B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment.
When we decreased $\omega$ from 0.3 to a conservative estimate of 0.15, 50-year extinction probability remained less than 0.5 for three populations and was 0.79 to 0.94 for three additional populations (@fig-viability B).
In contrast, for the six populations where yearly adult survival probability $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability C; population 70550, surveyed for 16 years).

@@

Disease-induced population declines are decimating global biodiversity [@daszak2000], but strategies to recover affected species are generally lacking.
Here, we tested the possibility that populations of resistant/tolerant individuals could be used to reestablish extirpated populations.
Although previously proposed [@mendelson2019; @scheele2021], this strategy has rarely been used in the wild (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (1) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (2) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog survival in the presence of Bd are fundamentally different between naive and recovering populations.
Following the arrival and establishment of Bd in previously-naive populations, adult frogs develop high Bd loads that lead to mass die-offs [@vredenburg2010].
In contrast, in recovering populations adult frogs typically have low-to-moderate and relatively constant Bd loads and mass die-offs are not observed [@briggs2010; @knapp2011, see also @fig-bdload-beforeafter].
The differences in Bd load of frogs from naive and recovering populations are also observed in controlled laboratory studies [@knapp2016, Figure 4].
The observed reduced susceptibility of frogs from recovering populations could in theory be due to several factors, including natural selection for more resistant/tolerant genotypes and/or acquired immunity, but until now evidence to evaluate the role of evolution was lacking.

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified specific alleles and genomic regions showing signatures of selection between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection (as measured by $\pi_{diff}$) contains the interferon-related gene GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014] and in populations differing in Bd susceptibility [@grogan2018].
This gene is also strongly linked to disease in salmon, explaining a notable 20% of the resistance phenotype [@robledo2020; @robledo2018].
We also identified a region, characterized by high *F~ST~* and low $\pi_{diff}$, that contained the complement genes C6 and C7.
This could indicate that balancing selection is acting in this region of the genome to favor a diverse set of alleles, as is known for C6 in humans [@soejima2005].
Finally, based on the analysis of individual outlier variants, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
This gene is associated with inflammation response and in *Xenopus* is highly expressed in wound stumps following amputation of tadpole tails [@fukazawa2009].
Also in this analysis, the outlier variant with the lowest p-value (uncharacterized gene LOC108802036) is located adjacent to a type I interferon gene (Np-IFNi2) in another frog genome [@gan2018], and together with GVINP1 further indicates the importance of interferon-related genes in this system.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering populations, individual and population-level impacts of Bd are still evident.
In an earlier study of 2 of our 12 translocated populations [@joseph2018], Bd infection and load had detectable effects on the survival of adults and may have influenced population establishment (sites referred to as "Alpine" and "Subalpine" in [@joseph2018] are identified as "70550" and "70505" in the current study).
Applying similar analyses to all 12 of our translocated populations would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on adults, the impacts on younger life stages are more apparent.
MYL frogs immediately following metamorphosis ("metamorphs") are highly susceptible to Bd infection [@ellison2018] and as a result experience high mortality.
This high susceptibility of metamorphs is documented for numerous species of anurans, and may result from the poorly developed immune system characteristic of this life stage [@humphries2022].
In naturally recovering and translocated MYL frog populations, we suggest that the high mortality of metamorphs is an important limitation on subsequent recruitment of new adults.
Therefore, although adult MYL frogs appear relatively resistant/tolerant, Bd infection continues to have important limiting effects on recovering populations [see also @hollanders2022].

@@

We collected DNA samples via buccal swabbing [@broquet2007] from 53 *Rana muscosa*/*Rana sierrae* individuals: 24 from four naive populations, and 29 from five recovering populations.
These populations were located in the southern Sierra Nevada, from northern Yosemite National Park to northern Sequoia National Park (@fig-allelefrequencies).
Samples were collected from 5-6 frogs per population.
To minimize potential confounding effects caused by variation in frog genotypes across latitude [e.g., @rothstein2020], we selected sampling sites such that both population types were represented across similar latitudinal ranges.
DNA was extracted following Qiagen DNEasy manufacturer's protocols.
We sequenced the samples using an exome capture assay as described in Byrne et al. (in review).
Briefly, genomic libraries were prepared and captured using a custom Nimblegen capture pool.
Capture baits were designed based on the coding regions of the *R. muscosa* transcriptome (GenBank accession GKCT00000000).
Captured libraries were then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the Vincent J. Coates Genomics Sequencing Lab at UC Berkeley.

@@

To visualize the genomic relationship of our populations we conducted a PCA using the glPCA function in the adegenet R package [@jombart2008].
To detect regions of the genome under selection between naive and recovering populations, we used two approaches: (1) a multivariate linear mixed model to evaluate individual variants (SNPs and INDELs), and (2) a splined window analysis to evaluate larger genomic regions.
For the variant analysis, we first used a stringent data filter to include only variants with \< 5% missing data (missing for no more than 2 individuals), and then calculated the likelihood ratio statistic for the resulting set of 148,307 high quality variants across 127 contigs using GEMMA [@zhou2014].
GEMMA calculates and incorporates a relatedness matrix for input samples, allowing us to account for relatedness and population structure when calculating likelihood ratio statistics.
We identified variants showing different allele frequencies between naive versus recovering populations ("outliers") using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot.
We developed a more liberal set of outlier variants using a Bonferroni- corrected significance level of 0.05 and used this set solely for the gene ontology (GO) analysis (see below and SI Appendix: Dataset S1, S2).

@@

Using the liberal set of variant outliers, we performed a statistical overrepresentation test (Fisher's exact test) using Blast2GO [@conesa2005].
First, we retrieved the BLAST hits and mapped GO terms for each gene in our targeted transcriptome.
Next, we compared the 30 unique outlier genes from the liberal set of outlier variants to the complete set of genes in our target transcriptome to determine if any GO biological functions, molecular functions, or cellular processes were overrepresented.
We then repeated this process for the set of 33 genes located in the overlapping *F~ST~* and $\pi_{diff}$ splined windows.
Code used for genomic analyses and to create figures is available from the following GitHub repository: <https://github.com/allie128/mylf-selection>.
<!--# Make sure that code has been uploaded to this link before manuscript submission. -->

@@

We conducted `r transno_min`--`r transno_max` translocations per site (@fig-yosemap) and each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`).
In preparation for each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed estimation of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and juvenile abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods.
For details of survey and swab analysis methods, see [@joseph2018].

>>>>

One consequence of human alteration of the biosphere is an increase in emerging infectious diseases [@jones2008; @fisher2012].
Such diseases pose a severe threat to wildlife populations [@daszak2000], and have caused dramatic declines and extinctions in a wide range of taxa, including echinoderms, mammals, birds, and amphibians [@hewson2014; @samuel2015; @scheele2019; @cunningham2021].
Amid this global loss of biodiversity, some populations decimated by disease have subsequently recovered to varying extents [@scheele2017; @voyles2018; @knapp2016].
Although we know relatively little about the mechanisms underlying recovery [@brannelly2021; @russell2020], host evolution is one possibility [@carlson2014; @searle2020].
This process is known from a diversity of wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], and often results in increased host resistance and/or tolerance [see @medzhitov2012 for definitions].
However, linking these evolutionary responses to population recoveries is difficult and few clear examples exist.
In addition to promoting recovery of disease-impacted populations, the resistant/tolerant individuals in such populations could potentially be used to reestablish populations that were extirpated following disease emergence [@mendelson2019; @brannelly2021].
To date, the use of resistant/tolerant individuals for species recovery is rare [but see @joseph2018], and studies that evaluate its potential utility are critically needed.

@@

Results from the individual variant and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C; SI Appendix: Datasets S1, S2).
One of the 7 identified genes (LOC108802036) did not have an associated annotation.
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong - and parallel - selection at this locus associated with exposure of frogs to Bd across multiple populations.

The splined window analysis identified 33 outlier regions for $\pi_{diff}$ and 58 outlier regions for *F~ST~* (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics ("shared regions") and 2 of these shared regions also contained one or more of the outlier SNPs described above.
A total of 33 annotated genes were found in the 9 shared regions (SI Appendix: Datasets S3, S4).
Given this large number of genes, here we focus on those with the strongest signal of selection and/or immune-related functions.
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a large negative $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended major histocompatibility complex (MHC) Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether frogs from recovering populations could be used to reestablish extirpated populations, we conducted `r n_distinct(d1$site_id, d1$date)` reintroductions in Yosemite National Park (`r min(d1$year)`--`r max(d1$year)`).
Each of the reintroductions involved collection of adult frogs from 1 of `r n_distinct(d1$donor)` recovering, Bd-positive "donor" populations and translocating them [@seddon2014] to 1 of `r n_distinct(d1$site_id)` nearby recipient sites (@fig-yosemap).
The donor populations included 2 of the 5 recovering populations in the frog evolution study described above (@fig-allelefrequencies A: population 1 and 4).
Following translocation, adult survival and recruitment of new adults was estimated from CMR surveys and counts of tadpoles and juveniles were obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1-year period following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate), which we suggest resulted primarily from our improved ability to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (@fig-bdload-beforeafter) [@joseph2018; @vredenburg2010].
Additional details on methods used to estimate Bd loads are provided in the SI Appendix.

Of the `r nrow(d2)` translocated populations, `r frog_counts_gt0_nrow` (`r frog_counts_gt0_nrow / nrow(d2)`) showed evidence of successful reproduction in subsequent years, as indicated by the presence of tadpoles and/or juveniles.
For these `r frog_counts_gt0_nrow` populations, one or both life stages were detected in nearly all survey-years following translocation (proportion of survey-years: median = `r frog_counts_gt0_median`, range = `r frog_counts_gt0_min`--`r frog_counts_gt0_max`).
These same populations were also those in which recruitment of new adults (i.e., progeny of translocated individuals) was detected.
As with early life stages, recruits were detected in the majority of post-translocation survey-years (proportion of survey-years: median = `r frog_captures_gt0_median`, range = `r frog_captures_gt0_min`--`r frog_captures_gt0_max`).
In summary, survey results indicate that translocations resulted in the establishment of reproducing MYL frog populations at most recipient sites despite the ongoing presence of Bd.

To determine whether Bd load had a negative effect on adult survival, as would be expected if frogs were highly susceptible to Bd infection, we developed linear and multilevel models.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (@fig-survival-postdens).
Important predictors included winter severity in the year following translocation, site elevation, and donor population (@fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (@fig-cond-effects, @fig-survival-postdens).
Collectively, these results are consistent with frogs in recovering populations having evolved sufficient resistance/tolerance to minimize the adverse effects of Bd infection.
In addition, the moderate Bd loads that characterized frogs following translocation (@fig-bdload-beforeafter) and relatively high survival of frogs in numerous translocated cohorts (@fig-translocation-survival) suggests that frogs translocated from recovering populations can maintain the benefits of resistance/tolerance in non-natal habitats.
Finally, in 3 locations where longer CMR time series allowed us to assess the survival of new adults recruited to the population, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs).
Therefore, frog resistance/tolerance in non-natal habitats is maintained across generations.

@@

Even when incorporating (i) demographic stochasticity and (ii) environmental stochasticity in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental variability in the presence of Bd), populations with high adult survival are likely to persist over a 50 year time horizon.
Our model predicted that, following a single introduction of 40 adult individuals into a population and using the baseline value for $\omega$ of 0.3, the six populations with the highest adult survival probabilities ($\sigma_{A_R} > 0.5$) had 50-year extinction probabilities of 0 to 0.32 (@fig-viability B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment.
When we decreased $\omega$ from 0.3 to a conservative estimate of 0.15, 50-year extinction probability remained less than 0.5 for three populations and was 0.79 to 0.94 for three additional populations (@fig-viability B).
In contrast, for the six populations where yearly adult survival probability $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability C; population 70550, surveyed for 16 years).

@@

Disease-induced population declines are decimating global biodiversity [@daszak2000], but strategies to recover affected species are generally lacking.
Here, we tested the possibility that populations of resistant/tolerant individuals could be used to reestablish extirpated populations.
Although previously proposed [@mendelson2019; @scheele2021], this strategy has rarely been used in the wild [see @joseph2018 for a small-scale example].
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (1) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease and that some of the genomic regions under selection contain genes related to disease resistance/tolerance, and (2) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog survival in the presence of Bd are fundamentally different between naive and recovering populations.
Following the arrival and establishment of Bd in previously-naive populations, adult frogs develop high Bd loads that lead to mass die-offs [@vredenburg2010].
In contrast, in recovering populations adult frogs typically have low-to-moderate and relatively constant Bd loads and mass die-offs are not observed [@briggs2010; @knapp2011, see also @fig-bdload-beforeafter].
The differences in Bd load of frogs from naive and recovering populations are also observed in controlled laboratory studies [see Figure 4 in @knapp2016].
The observed reduced susceptibility of frogs from recovering populations could in theory be due to several factors, including natural selection for more resistant/tolerant genotypes and/or acquired immunity, but until now evidence to evaluate the role of evolution was lacking.

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified specific alleles and genomic regions showing signatures of selection between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection (as measured by $\pi_{diff}$) contains the interferon-related gene GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014] and in populations differing in Bd susceptibility [@grogan2018].
This gene is also strongly linked to disease in salmon, explaining a notable 20% of the resistance phenotype [@robledo2020; @robledo2018].
We also identified a region, characterized by high *F~ST~* and low $\pi_{diff}$, that contained the complement genes C6 and C7.
The complement system plays an important role in innate immunity [@rieraromo2016], and our results could indicate that balancing selection is acting in this region of the genome to favor a diverse set of alleles, as is known for C6 in humans [@soejima2005].
Finally, based on the analysis of individual outlier variants, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
This gene is associated with inflammation response and in *Xenopus* is highly expressed in wound stumps following amputation of tadpole tails [@fukazawa2009].
Also in this analysis, the outlier variant with the lowest p-value (uncharacterized gene LOC108802036) is located adjacent to a type I interferon gene (Np-IFNi2) in another frog genome [@gan2018], and together with GVINP1 further suggests the importance of interferon-related genes in this system.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering populations, individual and population-level impacts of Bd are still evident.
In an earlier study of 2 of our 12 translocated populations [@joseph2018], Bd infection and load had detectable effects on the survival of adults and may have influenced population establishment (sites referred to as "Alpine" and "Subalpine" in @joseph2018 are identified as "70550" and "70505" in the current study).
Applying similar analyses to all 12 of our translocated populations would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on adults, the impacts on younger life stages are more apparent.
MYL frogs immediately following metamorphosis ("metamorphs") are highly susceptible to Bd infection [@ellison2018] and as a result experience high mortality.
This high susceptibility of metamorphs is documented in numerous species of anurans, and may result from the poorly developed immune system characteristic of this life stage [@humphries2022].
In naturally recovering and translocated MYL frog populations, we suggest that the high mortality of metamorphs is an important limitation on subsequent recruitment of new adults.
Therefore, although adult MYL frogs appear relatively resistant/tolerant, Bd infection continues to have important limiting effects on recovering populations [see also @hollanders2022].

@@

We collected DNA samples via buccal swabbing [@broquet2007] from 53 *Rana muscosa*/*Rana sierrae* individuals: 24 from 4 naive populations, and 29 from 5 recovering populations.
These populations are located in the southern Sierra Nevada, from northern Yosemite National Park to northern Sequoia National Park (@fig-allelefrequencies).
Samples were collected from 5-6 frogs per population.
To minimize potential confounding effects caused by variation in frog genotypes across latitude [e.g., @rothstein2020], we selected sampling sites such that both population types were represented across similar latitudinal ranges.
DNA was extracted following Qiagen DNEasy manufacturer's protocols.
We sequenced the samples using an exome capture assay as described in Byrne et al. (in review).
Briefly, genomic libraries were prepared and captured using a custom Nimblegen capture pool.
Capture baits were designed based on the coding regions of the *R. muscosa* transcriptome (GenBank accession GKCT00000000).
Captured libraries were then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the Vincent J. Coates Genomics Sequencing Lab at UC Berkeley.
Raw sequencing reads are available from NCBI SRA (PRJNA870451).

@@

To visualize the genomic relationships of our populations we conducted a PCA using the glPCA function in the adegenet R package [@jombart2008].
To detect regions of the genome under selection between naive and recovering populations, we used two approaches: (1) a multivariate linear mixed model to evaluate individual variants (SNPs and INDELs), and (2) a splined window analysis to evaluate larger genomic regions.
For the variant analysis, we first used a stringent data filter to include only variants with \< 5% missing data (missing for no more than 2 individuals), and then calculated the likelihood ratio statistic for the resulting set of 148,307 high quality variants across 127 contigs using GEMMA [@zhou2014].
GEMMA calculates and incorporates a relatedness matrix for input samples, allowing us to account for relatedness and population structure when calculating likelihood ratio statistics.
We identified variants showing different allele frequencies between naive versus recovering populations ("outliers") using a Bonferroni-corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot.
We developed a more liberal set of outlier variants using a Bonferroni-corrected significance level of 0.05 and used this set solely for the gene ontology (GO) analysis (see below and SI Appendix: Dataset S1, S2).

@@

Using the liberal set of outlier variants, we performed a statistical overrepresentation test (Fisher's exact test) using Blast2GO [@conesa2005].
First, we retrieved the BLAST hits and mapped GO terms for each gene in our targeted transcriptome.
Next, we compared the 30 unique outlier genes from the liberal set of outlier variants to the complete set of genes in our target transcriptome to determine if any GO biological functions, molecular functions, or cellular processes were overrepresented.
We then repeated this process for the set of 33 genes located in the overlapping *F~ST~* and $\pi_{diff}$ splined windows.
Code used for genomic analyses and to create figures is available from the following GitHub repository: <https://github.com/allie128/mylf-selection>.

@@

We conducted `r transno_min`--`r transno_max` translocations per site (@fig-yosemap, @fig-translocation-survival) and each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`).
In preparation for each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed estimation of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and juvenile abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
